A \texttt{shiny} New Opportunity for Interaction with Big Data in Undergraduate Education

### 2015-06-30 Body/Conclusions/Conclusions.Rnw
Major changes following committee discussion in defense.

\chapter{Conclusions}
\label{ThesisEnd}

@@

The study on traditional and simulation-based statistical inference curricula took an ambitious and statistically rigorous approach to comparing learning outcomes through a designed experiment. The random assignment of students to curriculum allows for the curriculum effect to be separated from the extraneous effects. The student learning outcomes, measured using the ARTIST scaled questions sets for confidence intervals and hypothesis tests, are analyzed using a bivariate MANCOVA model which controls for pre-treatment factors, including scores from a midterm and lab. The assumption of independence inherent in this model is critically assessed through a simulation study that eases the concern that a violation of the assumption would inflate the Type I error rate. The model found no significant curriculum effect for learning outcomes related to hypothesis tests; despite the added complexity in the simulation-based curriculum. The model does however indicate a statistically significant, 7\% improvement on the ARTIST scale for the simulation-based curriculum over the traditional curriculum for learning outcomes related to confidence intervals, while accounting for pre-treatment covariates. If this result is representative of true learning outcomes related to confidence intervals, then it is not only statistically significant, it is a finding that roars its affirmation of the advocates for simulation-based curriculum reform.

>>>>

\chapter{CONCLUSIONS}
\label{ThesisEnd}

@@

\km{The study on traditional and simulation-based statistical inference curricula took an ambitious and statistically rigorous approach to comparing learning outcomes through a designed experiment. The randomization of students to curricula helped to isolate the curricular effect by creating homogenous cohorts of students which received each curriculum. The student learning outcomes, measured using the ARTIST scaled questions sets for confidence intervals and hypothesis tests, were analyzed using a bivariate MANCOVA model which controls for pre-treatment factors, including scores from a midterm and lab. The model found no significant curriculum effect for learning outcomes related to hypothesis tests; despite the added complexity in the simulation-based curriculum. The model also indicates a statistically significant 7\% improvement on the ARTIST scale for the simulation-based curriculum over the traditional curriculum for learning outcomes related to confidence intervals, while accounting for pre-treatment covariates. However, the validity of these results is highly questionable because the model assumption of independence between students is likely untrue. The simulation study demonstrated that the Type I error rates in tests for curricular effects inflate dramatically under minor violations that have lecture or lab section based variance structure. Therefore, caution must be taken in interpreting the results of this study which are suggestive, but not conclusive, of a simulation-based curriculum benefit.}

### 2015-06-30 Body/ShinyDBSampler/ShinyDBSampler.Rnw
Major changes following committee discussion in defense.

\chapter{A \texttt{shiny} New Opportunity for Interaction with Big Data in Undergraduate Education}\label{ShinyDBSampler}

@@

\begin{center}
\textbf{Author}\\
Karsten Maurer, Iowa State University\\
\end{center}

\begin{center}
\textbf{Abstract}\\
\end{center}
As the availability of truly massive data proliferates, it is enticing to incorporate these data sets into the curriculum of an undergraduate statistics course.  Major barriers exist for interacting with big data due to the computationally intense nature of working with large databases.  Difficulties include gaining access to databases, interacting with database management software, and obtaining summary statistics or manageable subsamples from the database for student use.  This paper describes a web-based software application, the \textit{Shiny Database Sampler}, which allows instructors and students to bypass these barriers using a simple point-and-click interface constructed through \texttt{R} and the \texttt{R} packages \texttt{shiny} and \texttt{RMySQL}. The Shiny Database Sampler allows instructors and students to obtain subsamples from databases, using a variety of random sampling schemes. Application and evaluation of the software indicate that students find the interface easy to use, well connected to course concepts, and engaging through access to real data.

>>>>

\chapter{A SHINY NEW OPPORTUNITY FOR INTERACTION WITH BIG DATA IN UNDERGRADUATE EDUCATION}\label{ShinyDBSampler}

@@

\begin{center}
\textbf{Authors}\\
Karsten Maurer, Iowa State University, Primary Author\\
Heike Hofmann, Iowa State University\\
\end{center}

\begin{center}
\textbf{Abstract}\\
\end{center}

As the availability of truly massive data proliferates, it is enticing to incorporate these data sets into the curriculum of an undergraduate statistics course.  Major barriers exist for interacting with big data due to the computationally intense nature of working with large databases.  Difficulties include gaining access to databases, interacting with database management software, and obtaining summary statistics or manageable subsamples from the database for student use.  This paper describes a web-based software application, the \textit{Shiny Database Sampler}, which allows instructors and students to bypass these barriers using a simple point-and-click interface constructed through \texttt{R} and the \texttt{R} packages \texttt{shiny} and \texttt{RMySQL}. The Shiny Database Sampler allows instructors and students to obtain subsamples from databases, using a variety of random sampling schemes. Application and evaluation of the software indicate that students find the interface easy to use, well connected to course concepts, and engaging through access to real data.
